---
title: "Homework 8"
author: "Pranav Belligundu(psb898) - SDS 315 UT Austin"
date: "https://github.com/pranav-B21/SDS-315"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	fig.height = 5,
	fig.width = 8,
	warning = FALSE,
	tidy = TRUE,
	tidy.opts = list(width.cutoff = 60)
)
```


```{r echo = FALSE, message = FALSE, warnings = FALSE}
#libraries
library(ggplot2)
library(rvest)
library(tidyverse)
library(stringr)
library(kableExtra)
library(mosaic)
```

# **Problem 1: regression warm up**
### A. Load and examine the data
```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
creatinine <- read.csv('creatinine.csv')

model <- lm(creatclear ~ age, data = creatinine)

summary(model)

#A
predict(model, data.frame(age = 55))

#B
coef(model)["age"]

#C
pred_40 <- predict(model, data.frame(age = 40)) #predicted vals
pred_60 <- predict(model, data.frame(age = 60))
actual_40 <- 135 #actual vals
actual_60 <- 112

#diffs
diff_40 <- actual_40 - pred_40
diff_60 <- actual_60 - pred_60

diff_40
diff_60
```
### A
The expected value for a 55-year old is 133.72(mL/minute) clearance rate and I calcualted this through the lm function which gives me the intercept and the slope --> clearance rate = -0.6198159(age) + 147.81292. 

### B
For each increase in age by one, the creatine clearance rate decreases by -0.6198159 and this is the slope of the lin reg model.

### C
The difference in creatine rate for the 40 year old with a rate of 135 is 11.97 and the difference in creatine rate for the 60 year old with a rate of 112 is 1.38. Because the differnce of the 40 year old is higher, they are more healthier for their age. I determined this by using the model to predict it based on their ages and subtracted the actual rate to find the difference. 

# **Problem 2: Modeling disease growth**
```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
covid <- read.csv('covid.csv')

italy_data <- covid[covid$country == 'Italy',]

#model
covid_model_italy <- lm(log(deaths) ~ days_since_first_death, data = italy_data)
boot_covid_model_italy = do(10000)*lm(log(deaths) ~ days_since_first_death, data = resample(italy_data))
confint(boot_covid_model_italy, level = 0.95)

#growth rate
growth_rate_italy <- coef(covid_model_italy)["days_since_first_death"]

#doubling time
doubling_time_italy_l <- log(2) / 0.1586809 #lower
doubling_time_italy_u <- log(2) / 0.2078943 #upper

```
Italy: The growth rate with a 95% CI is between 0.159 and 0.208. The doubling time with a 95% CI is between 4.4 and 3.3 days. 


```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
spain_data <- covid[covid$country == 'Spain',]

#model
covid_model_spain <- lm(log(deaths) ~ days_since_first_death, data = spain_data)
boot_covid_model_spain = do(10000)*lm(log(deaths) ~ days_since_first_death, data = resample(spain_data))
confint(boot_covid_model_spain, level = 0.95)

#growth rate
growth_rate_spain <- coef(covid_model_spain)["days_since_first_death"]

#doubling time
doubling_time_spain_l <- log(2) / 0.2342065 #lower
doubling_time_spain_u <- log(2) / 0.3175116 #upper
```
Spain: The growth rate with a 95% CI is between 0.234 and 0.318. The doubling time with a 95% CI is between 3.0 and 2.2 days. 

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#line graph
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(
    title = "Daily Reported COVID-19 Deaths Over Time",
    x = "Days Since First Reported Death",
    y = "Number of Deaths",
    color = "Country"
  ) +
  theme_minimal()
```
# **Problem 3: price elasticity of demand**
```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
milk <- read.csv('milk.csv')

milk$log_price <- log(milk$price)
milk$log_sales <- log(milk$sales)

model <- lm(log_sales ~ log_price, data = milk)
boot_milk_model = do(10000)*lm(log_sales ~ log_price, data = resample(milk))
confint(boot_milk_model, level = 0.95)
```
The estimated price elasticity of demand for milk is −1.62, with a 95% bootstrapped confidence interval between −1.77 and −1.45. To estimate this, I log-transformed both price and sales to linearize the power-law demand model Q=KP^β, then used linear regression to estimate β as the slope of log_sales on log_price. I computed the confidence interval using 10,000 bootstrap resamples.

