---
title: "Homework 2"
author: "Pranav Belligundu(psb898) - SDS 315 UT Austin"
date: "https://github.com/pranav-B21/SDS-315/tree/main"
output:
  pdf_document:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```


```{r echo = FALSE, message = FALSE, warnings = FALSE}
#libraries
library(ggplot2)
library(rvest)
library(tidyverse)
library(stringr)
library(kableExtra)
```


# **Problem 1:** Beauty, or not, in the classroom

```{r echo = FALSE, message = FALSE, warnings = FALSE}
#reading the data for problem 1
profs <- read_csv('profs.csv', show_col_types = FALSE)
```


## Part A

```{r echo = FALSE, message = FALSE, warnings = FALSE, }
#Part A. Create a histogram to display the overall data distribution of course evaluation scores
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.1, fill = "grey", color = "black") +
  labs(title = "Distribution of Evaluation of Professor Ratings",
       x = "Evaluations",
       y = "Frequency")

#mean(profs$eval)
#median(profs$eval)
```
The x-axis represents the course evaluations scores and the y-axis is the frequency that a professor has gotten. The distribution for the course evaluations is skewed a little bit to the left, with the mean score being around 3.9 and the median being 4. 

## Part B

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ggplot(profs, aes(x = native, y = eval))+
  geom_boxplot() + 
  labs(title="Distribution of Course Evaluation based on if the Professor is a Native Speaker") +
  coord_flip() + 
  theme_minimal()
```
The x-axis represents if the professor was a native English speaker or not and the y-axis is the course evaluations score received. Based on the box plots, we can see that professors who are native English speakers received higher course evaluation scores than the professors that were non-native English speakers.                
     
## Part C

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.1, fill = "grey", color = "black") +
  labs(title = "Distribution of Professor Course Evaluations",
       x = "Course Evaluations",
       y = "Frequency") +
  facet_wrap(~gender) + 
  theme_minimal()
```

The x-axis represents the course evaluations scores and the y-axis is the frequency that a professor
has received a score in that range with the graphs faceted by male and female instructors. Based the histograms, we can see that male professors received higher course evaluation than female
professors. 

## Part D

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#Part D
ggplot(profs, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "grey") + # Add line of best fit
  labs(title = "Relationship between Professor’s Physical Attractiveness and Course Evaluations",
       x = "Physical Attractiveness",
       y = "Course Evaluations") +
  theme_minimal() +
  theme(legend.position = "none")
```
```{r echo = FALSE, message = FALSE, warnings = FALSE}
correlation <- cor(profs$beauty, profs$eval) 
```
The correlation between the professor’s physical attractiveness and course evaluations is `r cor(profs$beauty, profs$eval)` which is a small positive correlation. This shows that there is no correlation and there is no basis for an invdividual to believe that physical attractiveness does actually correlate with course evaluations.

# **Problem 2:** Bike Sharing

```{r echo = FALSE, message = FALSE, warnings = FALSE}
#reading the data for problem 2
bikes <- read_csv('bikeshare.csv', show_col_types = FALSE)
```

## Plot A

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#aggregate the data by hour and calculate the average total bike rentals
grouped_data <- group_by(bikes, hr)
hourly_averages <- summarise(grouped_data, average_rentals = mean(total))

#plot the data
ggplot(hourly_averages, aes(x = hr, y = average_rentals)) +
  geom_line() +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day",
       y = "Average Total Rentals") +
  theme_minimal()
```
The y-axis represents the counts for the average number of bike rentals and the x-axis represents each hour of the day. The graph shows clear peaks at around 8 AM and 5 PM which corresponds to normal commuting hours for those who go to work. This shows how bike rentals are driven by commuting patterns.

## Plot B

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#grouping the data
grouped_data <- group_by(bikes, hr, workingday)
hourly_averages_byworkingday <- summarise(grouped_data, average_rentals = mean(total))

#changing the names to convert number to character
labels <- setNames(c("Non-workday", "Workday"), c(0, 1))

#plot
ggplot(hourly_averages_byworkingday, aes(x = hr, y = average_rentals)) +
  geom_line() +
  facet_wrap(~workingday, ncol = 1, labeller = as_labeller(labels)) +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day",
       y = "Average Total Rentals") +
  theme_minimal()
```
The y-axis represents the counts for the average number of bike rentals and the x-axis represents each hour of the day which is faceted by working days (top) and non-working days (bottom). On working days, the rentals peak during commuting hours (8 AM and 5 PM). During non-working days, there is a small peak and decline around midday peak due to users using the rentals for fun.


## Plot C

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
bike_9 <- filter(bikes, hr == 9)
grouped_data <- group_by(bike_9, weathersit, workingday)
hourly_averages_byworkingday <- summarise(grouped_data, average_rentals = mean(total))

labels <- setNames(c("Non-workday", "Workday"), c(0, 1))

ggplot(hourly_averages_byworkingday, aes(x = as.factor(weathersit), y = average_rentals, fill = as.factor(weathersit))) +
  geom_bar(stat = "identity") +
  facet_wrap(~workingday, ncol = 1, labeller = as_labeller(labels)) +
  labs(title = "Average Ridership at 9 AM by Weather Situation",
       x = "Weather Situation Code",
       y = "Average Ridership") +
  scale_fill_discrete(name = "Weather Code") +
  theme_minimal()

```
This is a bar plot that shows the average number of bike rentals in the y-axis at 9 AM for different weather situations shown by the different colors. which is faceted by working days and non-working days. We can see that ridership decreases as weather conditions worsen and this effect is more evident during working days. This shows that weather significantly impacts morning commutes.

# **Problem 3:** Capital Metro UT Ridership

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#reading the data for problem 3
capmetro_UT <- read_csv('capmetro_UT.csv', show_col_types = FALSE)

capmetro_UT = mutate(capmetro_UT,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))

```

## Plot 1

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
capmetro_avg <- group_by(capmetro_UT, hour_of_day, day_of_week, month)
average_boardings <- summarise(capmetro_avg, average_boardings = mean(boarding))

ggplot(average_boardings, aes(x = hour_of_day, y = average_boardings, colour = as.factor(month))) +
  geom_line() +
  facet_wrap(~day_of_week) +
  labs(title = "Average Boardings by Hour, Day of Week, and Month",
       x = "Hour of the Day",
       y = "Average Boardings",
       colour = "Month") +
  theme_minimal()
```
This line graph shows the average boardings by hour of the day faceted by each day of the week with separate lines for each month. The hour of peak boardings is generally consistent occurring in the morning around 8 AM and evening around 5 PM. Lower average boardings during Mondays in September and Wed/Thu/Fri in November might indiciate holidays and days off.

## Plot 2

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ggplot(capmetro_UT, aes(x = temperature, y = boarding, colour = weekend)) +
  geom_point(size = 1, alpha = 0.5) +  # Adjust alpha for better visualization if points overlap
  facet_wrap(~hour_of_day) +
  labs(title = "Boardings vs. Temperature, Faceted by Hour of Day",
       x = "Temperature (°F)",
       y = "Boardings",
       colour = "Day Type") +
  theme_minimal()
```
This scatter plot shows the relationship between boardings and temperature, faceted by hour of the day.
Points are colored based on whether it is a weekend or a weekday where the red represents a weekday and blue represents a weekend. We can see that temperature appears to have a noticeable effect on boardings where a higher temperature correlates with higher ridership, especially during midday hours.


# **Problem 4:** Wrangling the Billboard Top 10

## Part A

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#reading the data for problem 3
billboards <- read_csv('billboard.csv', show_col_types = FALSE)
```

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
week_sub <- mutate(billboards, week=1)
top_ten <- group_by(week_sub, song, performer)
top_ten_weeks <- summarise(top_ten, total_weeks= sum(week))
top_ten_weeks <- arrange(top_ten_weeks, desc(total_weeks))
table <- head(top_ten_weeks,10)

table <- kable(table,col.names= c("Song", "Performer", "Total Weeks"),caption = "Top 10 Songs by Total Weeks on Billboard Top 100")

kable_styling(table,bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, position = "center")
```

This table above shows the top 10 most popular songs since 1958 which is ranked by the total number of weeks spent on the Billboard Top 100 chart. Columns include the song title, performer, and total count of weeks on the chart

## Part B

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
mus_diversity <- filter(billboards, year != 1958 & year != 2021)
mus_diversity <- group_by(mus_diversity,year)
mus_diversity <- summarise(mus_diversity, unique_song = length(unique(song)))

ggplot(mus_diversity) + 
  geom_line(aes(x=year,y=unique_song)) + 
  labs(title="Music Diversity by Year",
       x="Year",
       y="Musical Diversity") +
  theme_minimal()

```
The x-axis represents the year and the y-axis represents the number of unique songs. This line graph shows the musical diversity of the Billboard Top 100 from 1959 to 2020 wehre diversity is measured
by how many new, unique songs enter any position on the chart each year. 

## Part C

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ten_week_hit <- group_by(week_sub, song, performer)
ten_week_hit <- summarise(ten_week_hit, total_weeks= sum(week))
ten_week_hit <- filter(ten_week_hit, total_weeks >= 10)

ten_week_hit <- group_by(ten_week_hit,performer)
ten_week_hit <- summarise(ten_week_hit,songs=length(song))
ten_week_hit <- filter(ten_week_hit,songs >= 30)

ggplot(ten_week_hit) + geom_col(aes(fct_reorder(performer,songs),songs),fill="grey",color="black") + coord_flip() + theme_minimal() + labs(title="Artists with 30 Plus Ten Week Hits",y="Songs",x="Artists")

```

The x-axis represents the number of ten-weeks hits and the y-axis represents the artist. This bar graph shows 19 artists who have had at least 30 songs that appeared on the Billboard Top 100 for at least 10 weeks. 