---
title: "Homework 1"
author: "Pranav Belligundu - SDS 315 UT Austin"
date: "01-23-2025"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=4, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```


```{r echo = FALSE, message = FALSE, warnings = FALSE}
#libraries
library(ggplot2)
library(rvest)
library(tidyverse)
library(stringr)
library(kableExtra)
```

# **Problem 1:** Beauty, or not, in the classroom

### Part A
Create a histogram to display the overall data distribution of course evaluation scores

```{r echo = FALSE, message = FALSE, warnings = FALSE}
#reading the data for problem 1
profs <- read_csv('profs.csv', show_col_types = FALSE)
```

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#Part A. Create a histogram to display the overall data distribution of course evaluation scores
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black") +
  labs(title = "Distribution of Evaluation of Professor Ratings",
       x = "Eval",
       y = "Frequency")
```

### Part B
Use side-by-side boxplots to show the distribution of course evaluation scores by whether or not the professor is a native English speaker.

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ggplot(profs, aes(x = native, y = eval))+
  geom_boxplot() + 
  labs(title="Distribution of Course Evaluation based on if the Professor is a Native Speaker") +
  coord_flip() + 
  theme_minimal()
```

### Part C
Use a faceted histogram with two rows to compare the distribution of course evaluation scores for male #and female instructors.

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ggplot(profs, aes(x = eval)) +
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black") +
  labs(title = "Distribution of Professor Course Evaluations",
       x = "Course Evaluations",
       y = "Frequency") +
  facet_wrap(~gender) + 
  theme_minimal()
```

### Part D
Create a scatterplot to visualize the extent to which there may be an association between the professor’s physical attractiveness (x) and their course evaluations (y).

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#Part D
ggplot(profs, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + # Add line of best fit
  labs(title = "Relationship between Professor’s Physical Attractiveness and Course Evaluations",
       x = "Physical Attractiveness",
       y = "Course Evaluations") +
  theme_minimal() +
  theme(legend.position = "none")
```
```{r echo = FALSE, message = FALSE, warnings = FALSE}
correlation <- cor(profs$beauty, profs$eval) 
```
The correlation between the professor’s physical attractiveness and course evaluations is `r cor(profs$beauty, profs$eval)`. This shows that there is no correlation and there is no basis for an invdividual to believe that physical attractiveness does actually correlate with course evaluations.

# **Problem 2:** bike sharing

```{r echo = FALSE, message = FALSE, warnings = FALSE}
#reading the data for problem 2
bikes <- read_csv('bikeshare.csv', show_col_types = FALSE)
```

### Plot A
A line graph showing average hourly bike rentals (total) across all hours of the day (hr).

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#aggregate the data by hour and calculate the average total bike rentals
grouped_data <- group_by(bikes, hr)
hourly_averages <- summarise(grouped_data, average_rentals = mean(total))

#plot the data
ggplot(hourly_averages, aes(x = hr, y = average_rentals)) +
  geom_line() +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day",
       y = "Average Total Rentals") +
  theme_minimal()
```

### Plot B
A faceted line graph showing average bike rentals by hour of the day, faceted according to
whether it is a working day (workingday).

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#grouping the data
grouped_data <- group_by(bikes, hr, workingday)
hourly_averages_byworkingday <- summarise(grouped_data, average_rentals = mean(total))

#changing the names to convert number to character
labels <- setNames(c("Non-workday", "Workday"), c(0, 1))

#plot
ggplot(hourly_averages_byworkingday, aes(x = hr, y = average_rentals)) +
  geom_line() +
  facet_wrap(~workingday, ncol = 1, labeller = as_labeller(labels)) +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day",
       y = "Average Total Rentals") +
  theme_minimal()
```

### Plot C
A faceted bar plot showing average ridership (y) during the 9 AM hour by weather situation code (weathersit, x), faceted according to whether it is a working day or not. (Remember that you can focus on a specific subset of rows of a data set using filter.)

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
bike_9 <- filter(bikes, hr == 9)
grouped_data <- group_by(bike_9, weathersit, workingday)
hourly_averages_byworkingday <- summarise(grouped_data, average_rentals = mean(total))

labels <- setNames(c("Non-workday", "Workday"), c(0, 1))

ggplot(hourly_averages_byworkingday, aes(x = as.factor(weathersit), y = average_rentals, fill = as.factor(weathersit))) +
  geom_bar(stat = "identity") +
  facet_wrap(~workingday, ncol = 1, labeller = as_labeller(labels)) +
  labs(title = "Average Ridership at 9 AM by Weather Situation",
       x = "Weather Situation Code",
       y = "Average Ridership") +
  scale_fill_discrete(name = "Weather Code") +
  theme_minimal()

```

# **Problem 3:** Capital Metro UT Ridership

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#reading the data for problem 3
capmetro_UT <- read_csv('capmetro_UT.csv', show_col_types = FALSE)

capmetro_UT = mutate(capmetro_UT,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))

```

### Plot 1
One faceted line graph that plots average boardings by hour of the day, day of week, and month. You should facet by day of week. Each facet should include three lines of average boardings (y) by hour of the day (x), one line for each month and distinguished by color

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
capmetro_avg <- group_by(capmetro_UT, hour_of_day, day_of_week, month)
average_boardings <- summarise(capmetro_avg, average_boardings = mean(boarding))

ggplot(average_boardings, aes(x = hour_of_day, y = average_boardings, colour = as.factor(month))) +
  geom_line() +
  facet_wrap(~day_of_week) +
  labs(title = "Average Boardings by Hour, Day of Week, and Month",
       x = "Hour of the Day",
       y = "Average Boardings",
       colour = "Month") +
  theme_minimal()
```
### Plot 2
One faceted scatter plot showing boardings (y) vs. temperature (x), faceted by hour of the day, and
with points colored in according to whether it is a weekday or weekend

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ggplot(capmetro_UT, aes(x = temperature, y = boarding, colour = weekend)) +
  geom_point(size = 1, alpha = 0.5) +  # Adjust alpha for better visualization if points overlap
  facet_wrap(~hour_of_day) +
  labs(title = "Boardings vs. Temperature, Faceted by Hour of Day",
       x = "Temperature (°F)",
       y = "Boardings",
       colour = "Day Type") +
  theme_minimal()
```

# **Problem 4:** Wrangling the Billboard Top 10

## **Part. A**

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
#reading the data for problem 3
billboards <- read_csv('billboard.csv', show_col_types = FALSE)
```

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
week_sub <- mutate(billboards, week=1)
top_ten <- group_by(week_sub, song, performer)
top_ten_weeks <- summarise(top_ten, total_weeks= sum(week))
top_ten_weeks <- arrange(top_ten_weeks, desc(total_weeks))
table <- head(top_ten_weeks,10)

kable(table, format = "html", caption = "Top 10 Most Popular Songs on Billboard Top 100 Since 1958")
```
## **Part. B**

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
mus_diversity <- filter(billboards, year != 1958 & year != 2021)
mus_diversity <- group_by(mus_diversity,year)
mus_diversity <- summarise(mus_diversity, unique_song = length(unique(song)))

ggplot(mus_diversity) + 
  geom_line(aes(x=year,y=unique_song)) + 
  labs(title="Music Diversity by Year",
       x="Year",
       y="Musical Diversity") +
  theme_minimal()

```
## **Part. C**

```{r echo = FALSE, message = FALSE, warnings = FALSE, fig.width=8, fig.height=5}
ten_week_hit <- group_by(week_sub, song, performer)
ten_week_hit <- summarise(ten_week_hit, total_weeks= sum(week))
ten_week_hit <- filter(ten_week_hit, total_weeks >= 10)

ten_week_hit <- group_by(ten_week_hit,performer)
ten_week_hit <- summarise(ten_week_hit,songs=length(song))
ten_week_hit <- filter(ten_week_hit,songs >= 30)

ggplot(ten_week_hit) + geom_col(aes(fct_reorder(performer,songs),songs),fill="blue",color="black") + coord_flip() + theme_minimal() + labs(title="Artists with 30 Plus Ten Week Hits",y="Songs",x="Artists")

```